\documentclass{amsart}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{dsfont}
\usepackage{bbm}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{xifthen}
\usepackage{cancel}
\usepackage{hyperref}

\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-2in}
\calclayout

\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\HH}{\mathbb{H}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Zn}[1]{\mathbb{Z} / #1 \mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\dd}[2][ ]{\frac{\partial #1}{\partial #2}} % Partial derivative
\newcommand{\dsq}[3][ ]{\frac{\partial^2 #1}            % Second partial
{\ifthenelse { \equal {#2} {#3} }{\partial #2^2}{\partial #2 \partial #3}}}
\newcommand{\defarr}                        %definition iff arrow
{\overset{\textrm{def}}{\Longleftrightarrow}}
\newcommand{\defeq}                         %definition equality sign
{\overset{\textrm{def}}{=}}
\newcommand{\argeq}[1]                      %definition equality sign
{\overset{\textrm{#1}}{=}}

\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\orb}{orb}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\stab}{Stab}
\DeclareMathOperator{\sgn}{sign}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\tr}{tr}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}[thm]{Lemma}
\newtheorem*{unnumlemma}{Lemma}
\newtheorem{fact}[thm]{Fact}
\newtheorem{prop}[thm]{Propostion}
\newtheorem*{claim}{Claim}
\newtheorem{cor}{Corollary}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem*{rmk}{Remark}
\newtheorem*{ex}{Example}
%==============================================================================
% MATH 494 Collaborative Notes
% This is a collaborative notesheet that consists of notes from each day in Math
% 494, transcribed into Tex format for convenience and security from the
% Notebook Thief. General format:
% Contribute by adding a new section, titled "Month Day, Year"
% This is done in the format \section{Date}
% Tex notes using the numbering system used by Tasho himself. I've set up the
% counters on the theorems, definitions, and "facts" to run in tandem with the
% section that the content is in.
% Make sure to properly wrap all proofs, facts, and definitions using the above
% environments.
% Enjoy!
% Current Collaborators:
% 1. Pranav
% 2. Nick
% 3. Andrew
% 4. Ben
%==============================================================================
\begin{document}
\title{Math 494: Honors Algebra II}
\maketitle
\tableofcontents

\section{Jan. 27, 2017}

\begin{defn}
\hspace{0.5cm}
\begin{enumerate}
\item[(i)] A \textit{morphism of fields} is a morphism of rings whose source and target are fields
\item[(ii)] A \textit{subfield} is a subring of a field that is a field
\item[(iii)] An \textit{extension} of a field $k$ is a field $L$ containing $k$, denoted $L/k$.
\end{enumerate}
\end{defn}
 
\begin{rmk}
If $L/k$ is an extension, then $L$ is a $k$-algebra, and in particular a $k$-vector space. An extension $L/k$ is called \textit{finite} if $\dim_k(L) < \infty$. In this case define the \textit{degree} of $L/k$, denoted $[L : k] \stackrel{\text{def}}{=} \dim_k(L)$.\footnote{Debacker noted that the use of this notation, identical to the index of a subgroup, is because field extensions are in fact group-theoretic. When Galois groups are added to the notes, this will be discussed below.}
 \end{rmk}
 
\begin{fact}
 If $k_1, k_2$ are subgields of $L$, $k_1 \cap k_2$ is a subfield.
\end{fact}

\begin{defn}
\hspace{0.5cm}
\begin{enumerate}
\item[(i)] Every field has a smallest field, called the \textit{prime subfield}, equal to $\bigcap_{k \subset L} k^{\text{field}}$, the intersection of all subfields of $L$. \\
\textit{Example:} The prime subfield of $\RR$ is $\QQ$. To see this, observe: any subfield of $\RR$ must contain 1 and 0, thus sums of the form $1 + \cdots + 1$, so it must contain $\NN$. Throwing in additive and then multiplicative inverses gives $\QQ$.
\item[(ii)] Given an extension $L/k$ and any subset $S \subset L$, let $k(S)$ be the smallest subfield of $L$ containing $k$ and $S$. 
\end{enumerate}
\end{defn}

\begin{prop}
Let $L$ be a ring and $k \subset L$ a subfield. If $L$ is a domain and $\dim_k(L) < \infty$ then $L$ is a field.
\end{prop}
\begin{proof}
Let $0 \neq a \in L$. The map $a \cdot \colon L \rightarrow L$ that sends $x$ to $ax$ is $k$-linear, and since $L$ is a domain, injective: if $ax = 0$ then since $a \neq 0$, $x = 0$. By rank-nullity, the dimension of the image of $a \cdot$ must be equal to that of the target, i.e. $a \cdot$ is surjective. Thus there exists $G \in L$ so that $aG = 1$, implying that, as desired, $L$ is a field.
\end{proof}

\begin{defn}
Let $L/k$ be an extension, $L \supset E, F \supset k$ intermediate fields, each finite over $k$. The \textit{composite of E and F} is 
\[E \cdot F = \left\{ \sum_{i = 1}^n e_i \cdot f_i \colon n \in \NN, e_i \in E, f_i \in F \right\}.\]
\end{defn}

\begin{prop}
$EF$ is a field extension of $k$. It is finite and $[EF : k] \leq [E : k] \cdot [F : k]$.
\end{prop}
\begin{proof}
It is clear that $EF$ is a subring of $L$ containing $k$. Let $\{e_1, ..., e_n\}$ and $\{f_1, ..., f_m\}$ be, respectively, bases for $E/k$ and $F/k$. Then the set $\{e_i \cdot f_j \colon i \in \{1, ..., n\}, j \in \{1, ..., m\}\}$ spans the $k$-vector space $EF$. Thus by elementary results, $\dim_k(EF) \leq \#\{e_i \cdot f_j\} = [E : k] \cdot [F : k]$. In particular, $[EF : k] < \infty$. Applying \textbf{Proposition 1.4} above, we have that $EF$ is a field.
\end{proof}

\begin{prop}
Let $L/k$ be a finite extension, $V$ a finite dimensional $L$-vector space. Then $V$ is finite dimensional as a $k$-vector space, with $\dim_k(V) = \dim_L(V) \cdot [L : k]$.
\end{prop}
\begin{proof}
Let $\{v_1, ..., v_n\}$ be a basis for $V$ over $L$, $\{\ell_1, ..., \ell_m\}$ be a basis for $L$ over $k$. We will show $\{\ell_i \cdot v_j \colon i \in \{1, ..., n\}, j \in \{1, ..., j\}\}$ is a basis for $V$ over $k$:
\begin{enumerate}
\item[(i)] Spanning: given $x \in V$ we may write $x \in \lambda_1v_1 + \cdots + \lambda_nv_n$ for $\lambda_i \in L$, with each $\lambda_i = \mu_{i,1}\ell_1 + \cdots + \mu_{i,m}\ell_m$. Then $x = \sum_i \sum_j \mu_{i, j} \ell_j v_i$ with the $\mu_{i,j} \in k$.
\item[(ii)] Linear independence: Write $\sum \mu_{i,j} \ell_j v_i = 0$. Then $0 = \sum_i \left( \sum_j\mu_{i,j} \ell_j \right) v_i$. By linear independence of the $v_i$ we have $\sum_j\mu_{i,j} \ell_j = 0$, by linear independence of the $\ell_j$ we have $\mu_{i,j} = 0$ for all $i,j$.
\end{enumerate}
\end{proof}

\begin{cor}
If $M/L/K$ is a tower of finite ixtensions, then $[M : k] = [M : L] \cdot [L : k]$.
\end{cor}

\begin{defn}
Let $L/k$ be an extension, and $a \in L$. If there is $0 \neq p \in k[T]$ such that $p(a) = 0$, then $a$ is \textit{algebraic over $k$}. Otherwise, $a$ is \textit{transcendental over $k$}.
\end{defn}

As per the above definitions, we would like to study the following construction: let $\text{ev} \colon k[T] \rightarrow L$, $p \mapsto p(a)$ be the evaluation morphism. If $a$ is transcendental, ev is injective. Further, since $L$ is a field, we have that ev factors uniquely through the fraction field $k(T)$\footnote{This notation was established on homework: $k(T)$ is the set of rational functions with coefficients in $k$.} and induces an isomorphism $k(T) \stackrel{\sim}{\rightarrow} k(a)$. If $a$ is algebraic, let $0 \neq I \subset k[T]$ be the kernel of ev. Since $k[T]$ is a Euclidean domain, hence a PID (see a proposition from Jan. 20), so there exists a unique monic polynomial $p \in k[T]$ such that $(p) = I$.

\begin{defn}
$p$ is called the \textit{minimal polynomial} of $a$ over $k$ $a/k$, written $Ma/k$. $deg(a/k) := \deg(Ma/k)$ is its \textit{degree}.
\end{defn}

\begin{rmk}
$Ma/k$ is also the minimal polynomial of the endomorphism (linear operator that is not necessarily an isomorphism) $a \cdot \colon L \rightarrow L$.
\end{rmk}
\end{document}